Metrics

### 2023-11-17 src/metrics.Rmd
Calculated metrics

```{r, echo=FALSE}
knitr::opts_chunk$set(
  message=FALSE,
  warning=FALSE
)
```

@@

```{r}
library(tidyverse)

options(readr.show_col_types = FALSE)
```

@@

    # Calc N num of causes if not known
    N <- if (is.null(N)) length(unique(c(actual, pred))) else N

    # Calc frac of deaths in top k causes
    C <- sum(actual == pred) / len(actual)

@@

Load data from `data` and `tmp` folders.

```{r}
df <- list(
    wva2016_icd10 = read_csv(
        "../data/wva2016_icd10_v1.csv",
        col_types = cols(wva2016_code = col_double())
    ),
    icd10_cghr10 = read_csv("../data/icd10_cghr10_v1.csv"),
    adult = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_adult_v1.csv",
            col_select = "rowid"
        ) %>% mutate(round = 1, age = "adult"),
        read_csv(
            "../tmp/healsl_rd2_adult_v1.csv",
            col_select = "rowid"
        ) %>% mutate(round = 2, age = "adult"),
    ),
    child = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_child_v1.csv",
            col_select = "rowid"
        ) %>% mutate(round = 1, age = "child"),
        read_csv(
            "../tmp/healsl_rd2_child_v1.csv",
            col_select = "rowid"
        ) %>% mutate(round = 2, age = "child")
    ),
    neo = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_neo_v1.csv",
            col_select = "rowid"
        ) %>% mutate(round = 1, age = "neo"),
        read_csv(
            "../tmp/healsl_rd2_neo_v1.csv",
            col_select = "rowid"
        ) %>% mutate(round = 2, age = "neo")
    ),
    model_chatgpt3 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_chatgpt3_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_chatgpt3_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        )
    ),
    model_chatgpt4 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_chatgpt4_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_chatgpt4_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        )
    ),
    model_insilicova = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_insilicova_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_insilicova_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        )
    ),
    model_interva5 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_interva5_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_interva5_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        )
    )
)
```

```{r, echo=FALSE}
for (i in names(df)) {
    print(i)
    print(head(df[[i]]))
}
```

## Preprocess

Convert ICD-10 and WVA-2016 codes to CGHR-10 codes.

```{r}

# Combine all models
models <- df[str_starts(names(df), "model_")]

# Add age groups to predictions
df$chatgpt3 <- df$chatgpt3 %>%
    left_join(
        df$adult,
        by = c(
            "rowid" = "rowid",
            "round" = "round"
        ),
        multiple = "first"
    )

# Convert wva2016 to icd10 via join
df$insilicova <- df$insilicova %>%
    left_join(
        df$wva2016_icd10,
        by = c("cause1_wva" = "wva2016_code"),
        multiple = "first"
    )
df$interva5 <- df$interva5 %>%
    left_join(
        df$wva2016_icd10,
        by = c("cause1_wva" = "wva2016_code"),
        multiple = "first"
    )

# Convert icd10 to cghr10 via join
df$chatgpt3 <- df$chatgpt3 %>%
    left_join(
        df$icd10_cghr10,
        by = c("api_chat_icd" = "icd10_code"),
        multiple = "first"
    )
df$interva5 <- df$interva5 %>%
    left_join(
        df$wva2016_icd10,
        by = c("cause1_wva" = "wva2016_code"),
        multiple = "first"
    )
```

>>>>

```{r, echo=FALSE}
knitr::opts_chunk$set(
  message=FALSE,
  warning=FALSE
)
options(readr.show_col_types = FALSE)
```

@@

```{r}
library(tidyverse)
```

@@

    # Calc N num of causes if not known
    N <- if (is.null(N)) length(unique(na.omit(c(actual, pred)))) else N

    # Calc frac of deaths in top k causes
    TP <- actual == pred
    TP[is.na(TP) | is.null(TP)] <- FALSE # for no preds
    C <- sum(TP) / length(actual)

@@

Load data from `data` folder.

```{r}
df <- read_csv("../data/healsl_rd1to2_cod_v1.csv")
```

```{r, echo=FALSE}

# Display num of cases
cat("Number of Cases: ", nrow(df), "\n")

# Get actual and pred
actual <- df$physician_cghr10
pred <- df %>% select(ends_with("_cghr10"), -physician_cghr10)

# Physician info
cat(paste0(
    "Physician Coded Cases (",
    length(unique(na.omit(actual))), " CODs): ",
    length(actual) - sum(is.na(actual) | is.null(actual)),
    "\n"
))

# Models info
for (mcol in colnames(pred)) {
    cat(paste0(
        toupper(str_remove(mcol, "_cghr10")), " Predicted Cases (", 
        length(unique(na.omit(pred[[mcol]]))), " CODs): ",
        length(pred[[mcol]]) - sum(is.na(pred[[mcol]]) | is.null(pred[[mcol]])),
        "\n"
    ))
}

print(head(df))
```

Filter for cases that were coded by physicians.

```{r}
df <- df %>%
    filter(!is.na(physician_cghr10) & !is.null(physician_cghr10))
```

```{r, echo=FALSE}
cat("Number of Cases for Metrics: ", nrow(df))
print(head(df))
```

Create model combinations where if any of the models have the physician code, then set the combined model's output for the case to be the physician code.

```{r}
df <- df %>%
    mutate(
        "chatgpt3_insilicova_cghr10" = if_else(
            chatgpt3_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt3_cghr10
        ),
        "chatgpt4_insilicova_cghr10" = if_else(
            chatgpt4_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt4_cghr10
        ),
        "chatgpt3_interva5_cghr10" = if_else(
            chatgpt3_cghr10 == physician_cghr10 | 
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt3_cghr10
        ),
        "chatgpt4_interva5_cghr10" = if_else(
            chatgpt4_cghr10 == physician_cghr10 | 
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt4_cghr10
        ),
        "chatgpt3_insilicova_interva5_cghr10" = if_else(
            chatgpt3_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10 |
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt3_cghr10
        ),
        "chatgpt4_insilicova_interva5_cghr10" = if_else(
            chatgpt4_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10 |
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            chatgpt4_cghr10
        )
    )
```

```{r, echo=FALSE}
print(head(df %>% select(
    chatgpt3_insilicova_cghr10,
    chatgpt4_insilicova_cghr10,
    chatgpt3_interva5_cghr10,
    chatgpt4_interva5_cghr10,
    chatgpt3_insilicova_interva5_cghr10,
    chatgpt4_insilicova_interva5_cghr10
)))
```

Separate physician (actual) and model (predicted) CODs.

```{r}
actual <- df$physician_cghr10
pred <- df %>%
    select(ends_with("_cghr10"), -physician_cghr10) %>%
    rename_all(~toupper(str_replace(., "_cghr10", "")))
```

```{r, echo=FALSE}
cat(paste0(
    "Physician Coded Cases (",
    length(unique(na.omit(actual))), " CODs): ",
    length(actual) - sum(is.na(actual) | is.null(actual)),
    "\n"
))
```

```{r, echo=FALSE}
for (mcol in colnames(pred)) {
    cat(paste0(
        mcol, " Predicted Cases (", 
        length(unique(na.omit(pred[[mcol]]))), " CODs): ",
        length(pred[[mcol]]) - sum(is.na(pred[[mcol]]) | is.null(pred[[mcol]])),
        "\n"
    ))
}
```

## Methods

### PCCC

Calculate PCCC to evaluate indivudal performance for each model.

```{r}
pccc <- pred %>%
    pivot_longer(
        everything(),
        names_to = "Model",
        values_to = "prediction"
    ) %>%
    group_by(Model) %>%
    summarise(
        PCCC = calc_pccc(actual, prediction, k = str_count(unique(Model), "_") + 1)
    ) %>%
    mutate(Model = str_replace_all(Model, "_", " & ")) %>%
    arrange(desc(PCCC))
pccc
```

### CSMF Accuracy

Calculate CSMF Accuracy to evaluate population performance for each model.

```{r}
csmf_acc <- pred %>%
    pivot_longer(
        everything(),
        names_to = "Model",
        values_to = "prediction"
    ) %>%
    group_by(Model) %>%
    summarise(
        "CSMF Accuracy" = calc_csmf_acc(actual, prediction)
    ) %>%
    mutate(Model = str_replace_all(Model, "_", " & ")) %>%
    arrange(desc(`CSMF Accuracy`))
csmf_acc
```
